Project Creditworthiness

### 2021-07-30 Project 3/Assignment/Creditworthiness/Submission - NND.Rmd
Revise submission file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

@@

###	What data is needed to inform those decisions?
Outcome: Credit-Application-Result.

@@

```{r import, message=c(32), warning=FALSE}
library(tidyverse)
library(caret)
library(randomForest)
library(bst)
library(cowplot)
#raw <- read.csv("./Creditworthiness/credit-data-training.csv", stringsAsFactors=F)
#validation <- read.csv("./Creditworthiness/customers-to-score.csv", stringsAsFactors=F)
colnames(validation) <- str_replace_all(colnames(validation), '\\.', '_')
colnames(raw) <- str_replace_all(colnames(raw), '\\.', '_')
glimpse(raw)

@@

### Confusion matrices
```{r training}
# Logistic regression
lr <- vector()
model1 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method='glm')  #check
y_hat1 <- predict(model1, test)
y_hat10 <- predict(model1, train)
lr['Accuracy_train'] <- confusionMatrix(data = y_hat10, reference = train$Credit_Application_Result)$overall["Accuracy"] #Bias
lr['Accuracy_test'] <- confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result)$overall["Accuracy"]

model1w <- train(worthy~., 
                  data=train[!colnames(train) %in% c('Credit_Application_Result', 'nonworthy')], 
                  method='glm')
y_hat1w <- predict(model1w, test)
lr['Accyracy_worthy'] <- confusionMatrix(data = y_hat1w, reference = test$worthy)$overall["Accuracy"]

@@

# Decision tree
tree <- vector()
model2 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')],
                method='rpart')
y_hat2 <- predict(model2, test)
y_hat20 <- predict(model2, train)
tree['Accuracy_train'] <- confusionMatrix(data = y_hat20, reference = train$Credit_Application_Result)$overall["Accuracy"] 
tree['Accuracy_test'] <- confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result)$overall["Accuracy"]

@@

# Forest model
forest <- vector()
set.seed(1)
model3 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method='rf',
                ntree=500)
y_hat3 <- predict(model3, test)
y_hat30 <- predict(model3, train)
forest['Accuracy_train'] <- confusionMatrix(data = y_hat30, reference = train$Credit_Application_Result)$overall["Accuracy"]
forest['Accuracy_test'] <- confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result)$overall["Accuracy"]

@@

test4 <- test %>% mutate(#Credit_Application_Result= as.numeric(Credit_Application_Result),
  Account_Balance= as.factor(as.numeric(Account_Balance)),
  Payment_Status_of_Previous_Credit=as.factor(as.numeric(Payment_Status_of_Previous_Credit)),
  Purpose=as.factor(as.numeric(Purpose)),
  Value_Savings_Stocks=as.factor(as.numeric(Value_Savings_Stocks)),
  Length_of_current_employment=as.factor(as.numeric(Length_of_current_employment)),
  No_of_Credits_at_this_Bank=as.factor(as.numeric(No_of_Credits_at_this_Bank)),
  )
control <- trainControl(method = "cv", number = 5)
set.seed(1)
model4 <- train(Credit_Application_Result~., 
                data=train4[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method="bstTree", 
                trControl = control)
y_hat4 <- predict(model4, test4)
y_hat40 <- predict(model4, train4)
boosted['Accuracy_train'] <- confusionMatrix(data = y_hat40, reference = train4$Credit_Application_Result)$overall["Accuracy"]  
boosted['Accuracy_test'] <- confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result)$overall["Accuracy"]

@@

#Print confusion matrices
print(confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result))
```
### Overall accuracy
Among the 4 models, Logistic regression has the highest overall accuracy agains the Validation set (0.83). The bias can be seen in Logistic regression, Random forest, and Boosted tree, since accuracy against train set is lower than validation test for those models. 
```{r acc train test}
print(data.frame(lr, tree, forest, boosted)[c('Accuracy_train','Accuracy_test'),])
```
### Plot variable importance
Most important predictors:
- For Logistic regression: Account_Balance, Payment_Status_of_Previous_Credit, Credit_Amoun, Value_Savings_Stocks, Purpose, Instalment_per_cent.
- For Decision tree: Payment_Status_of_Previous_Credit, Credit_Amount, Value_Savings_Stocks, Account_Balance, Duration_of_Credit_Month.
- For Random forest: Credit_Amount, Age_years, Duration_of_Credit_Month.
- For Boosted tree: Account_Balance, Value_Savings_Stocks, Paymen_Status_of_Previous_Credit, and Duration_of_Credit_Month are the 4 most important variables. The difference is not remarkable among the fields.
```{r plot1}
plot1
plot2
plot3
plot4
```

## Step 4: Writeup
Accuracy table
```{r acc}
print(data.frame(lr, tree, forest, boosted))
```

@@

```
Logistic regresson model has the highest overall accuracy against the Validation set, as well as the highest accuracy within segments. It is also has the most optimal ROC. Therefore, I choose to use logistic regression model.
Applying the chosen model, the number of individuals are creditworthy is 413.
```{r predict}
yhat_final <- predict(model1, validation)
print(sum(yhat_final=='Creditworthy'))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error=TRUE)
```

@@

###	What data is needed to inform those decisions?
- Data set of past applications for training models:
  - Outcome: Credit-Application-Result.
  - Potention predictors: Duration_of_Credit_Month, Credit_Amount, Age_years, Instalment_per_cent, Most_valuable_available_asset, Type_of_apartment, and so on.
- New data set (without outcome variable) to apply the chosen model and make decisions.

@@

```{r import, message=c(32), warning=FALSE}
library(tidyverse)
library(caret)
library(randomForest)
library(bst)
library(cowplot)
library(pROC)
#raw <- read.csv("./Creditworthiness/credit-data-training.csv", stringsAsFactors=F)
#validation <- read.csv("./Creditworthiness/customers-to-score.csv", stringsAsFactors=F)
colnames(validation) <- str_replace_all(colnames(validation), '\\.', '_')
colnames(raw) <- str_replace_all(colnames(raw), '\\.', '_')
glimpse(raw)

@@

### Confusion matrices
```{r training}
# Logistic regression
lr <- vector()
model1 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method='glm')  #check
y_hat1 <- predict(model1, test)
y_hat10 <- predict(model1, train)
lr['Accuracy_train'] <- confusionMatrix(data = y_hat10, reference = train$Credit_Application_Result)$overall["Accuracy"] #Bias
lr['Accuracy_test'] <- confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result)$overall["Accuracy"]
lr['PPV'] <- confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result)$byClass['Pos Pred Value']
lr['NPV'] <- confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result)$byClass['Neg Pred Value']

model1w <- train(worthy~., 
                  data=train[!colnames(train) %in% c('Credit_Application_Result', 'nonworthy')], 
                  method='glm')
y_hat1w <- predict(model1w, test)
lr['Accuracy_worthy'] <- confusionMatrix(data = y_hat1w, reference = test$worthy)$overall["Accuracy"]

@@

# Decision tree
tree <- vector()
model2 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')],
                method='rpart')
y_hat2 <- predict(model2, test)
y_hat20 <- predict(model2, train)
tree['Accuracy_train'] <- confusionMatrix(data = y_hat20, reference = train$Credit_Application_Result)$overall["Accuracy"] 
tree['Accuracy_test'] <- confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result)$overall["Accuracy"]            
tree['PPV'] <- confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result)$byClass['Pos Pred Value']
tree['NPV'] <- confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result)$byClass['Neg Pred Value']

@@

# Forest model
forest <- vector()
set.seed(1)
model3 <- train(Credit_Application_Result~., 
                data=train[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method='rf',
                ntree=500)
y_hat3 <- predict(model3, test)
y_hat30 <- predict(model3, train)
forest['Accuracy_train'] <- confusionMatrix(data = y_hat30, reference = train$Credit_Application_Result)$overall["Accuracy"]
forest['Accuracy_test'] <- confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result)$overall["Accuracy"]   
forest['PPV'] <- confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result)$byClass['Pos Pred Value']
forest['NPV'] <- confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result)$byClass['Neg Pred Value']

@@

test4 <- test %>% mutate(#Credit_Application_Result= as.numeric(Credit_Application_Result),
  Account_Balance= as.factor(as.numeric(Account_Balance)),
  Payment_Status_of_Previous_Credit=as.factor(as.numeric(Payment_Status_of_Previous_Credit)),
  Purpose=as.factor(as.numeric(Purpose)),
  Value_Savings_Stocks=as.factor(as.numeric(Value_Savings_Stocks)),
  Length_of_current_employment=as.factor(as.numeric(Length_of_current_employment)),
  No_of_Credits_at_this_Bank=as.factor(as.numeric(No_of_Credits_at_this_Bank)),
  )
control <- trainControl(method = "cv", number = 5)
set.seed(1)
model4 <- train(Credit_Application_Result~., 
                data=train4[!colnames(train) %in% c('worthy', 'nonworthy')], 
                method="bstTree", 
                trControl = control)
y_hat4 <- predict(model4, test4)
y_hat40 <- predict(model4, train4)
boosted['Accuracy_train'] <- confusionMatrix(data = y_hat40, reference = train4$Credit_Application_Result)$overall["Accuracy"]  
boosted['Accuracy_test'] <- confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result)$overall["Accuracy"]  
boosted['PPV'] <- confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result)$byClass['Pos Pred Value']
boosted['NPV'] <- confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result)$byClass['Neg Pred Value']

@@

#Print confusion matrices
print(confusionMatrix(data = y_hat1, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat2, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat3, reference = test$Credit_Application_Result))
print(confusionMatrix(data = y_hat4, reference = test4$Credit_Application_Result))
```
### Overall accuracy
Among the 4 models, Logistic regression has the highest overall accuracy agains the Validation set (0.83). The bias was not observed in Logistic regression, Random forest, and Boosted tree model considering PPV and NPV. 
```{r acc train test}
print(data.frame(lr, tree, forest, boosted)[c('Accuracy_train','Accuracy_test','PPV','NPV'),])
```
### Plot variable importance
Most important predictors:
- For Logistic regression: Account_Balance, Payment_Status_of_Previous_Credit, Credit_Amoun, Value_Savings_Stocks, Purpose, Instalment_per_cent.
- For Decision tree: Payment_Status_of_Previous_Credit, Credit_Amount, Value_Savings_Stocks, Account_Balance, Duration_of_Credit_Month.
- For Random forest: Credit_Amount, Age_years, Duration_of_Credit_Month.
- For Boosted tree: Account_Balance, Value_Savings_Stocks, Paymen_Status_of_Previous_Credit, and Duration_of_Credit_Month are the 4 most important variables. The difference is not remarkable among the fields.
```{r plot1}
plot1
plot2
plot3
plot4
```

## Step 4: Writeup
Accuracy table
```{r acc}
print(data.frame(lr, tree, forest, boosted)[c('Accuracy_train','Accuracy_test','PPV','NPV'),])
```

@@

```
Logistic regresson model is not bias and has the highest overall accuracy against the Validation set. It is also has the most optimal ROC. Therefore, I choose to use logistic regression model.
Applying the chosen model, the number of individuals are creditworthy is 413.
```{r predict}
yhat_final <- predict(model1, validation)
print(sum(yhat_final=='Creditworthy'))
```
